A spiral galaxy's mass distribution uncovered through lensing and dynamics

### 2015-09-04 docs/discussion_ML_v1.01.tex
Finished section on mass / light comparison.

\subsection{Stellar mass-to-light ratios}

Further knowledge of the stellar populations within J1331 and their initial mass functions (IMF) would be useful. In particular a sophisticated guess for the \emph{stellar} mass-to-light ratio in J1331's bulge could be compared to our very reliable measurement of the \emph{total} mass-to-light ratio inside the Einstein radius (see Table \ref{tab:bestfitlensmodel}) to support or contradict the presence of a lot of dark matter in the bulge. Traditional choices for the IMF are the bottom-heavy IMF by \citet{Salpeter1955},

@@

\\We compare our findings of the stellar mass-to-light ratios with the study by \cite{SWELLSV}: \cite{SWELLSV} found that the bulge of J1331 has an IMF more bottom-heavy than the Salpeter IMF from fitting a NFW halo to (1) the Einstein mass and (2) gas kinematics at larger radii $\gtrsim 8$ arcsec. We fitted the stellar kinematics within $\sim 3.5$ arcsec and the Einstein mass with a NFW halo and found $\Upsilon^\text{dyn}_{*,I} = 4.2 \pm 0.2$ (see Table \ref{tab:modelB4_bestfit}). \cite{SWELLSV}  found systematically lower NFW halos ($v_\text{circ}(5\text{arcsec}) \sim 120 \text{km s}^{-1}$ according to their Figure 2) than we did ($v_\text{circ}(5\text{arcsec}) \sim 200 \text{km s}^{-1}$, see Figure \ref{fig:kinematics}). [TO DO: Continue writing notes into text]

@@

\item We found an Einstein mass-to-light ratio of $\Upsilon_\text{I,tot,Ein} = 5.6$ (see Table \ref{tab:bestfitlensmodel}). In case of a lower mass NFW halo, the largest contribution to this mass-to-light ratio would then be stellar mass. And the corresponding IMF would be more bottom heavy than Salpeter as well, like what \citet{SWELLSV} found.
\item The light distribution multiplied with $\Upsilon_\text{I,tot,Ein}$ looks very similar to \cite{SWELLSV} [TO DO: Look details up in their paper and my work]

@@

\item The above discussion motivates the following speculation: In the absence of a strong dark matter contribution in the center, the overall stellar-mass-to-light ratio within the Einstein radius indicates a bottom-heavy Salpeter-like IMF, consistent with estimates from the velocity dispersion. At the same time the central dip can then be only explained, if there was an increase in stellar mass-to-light ratio with radius \textit{within} the bulge. \emph{If} there was a stellar population in the very center with an IMF close to the \citet{Chabrier2003} IMF, surrounded by a more bottom-heavy population, the central stellar kinematics would be well explained and be fully consistent with the lensing results, as found in Figure \ref{fig:modelG} and the surrounding text. It is not expected that the bluish disk of J1331 has a $\Upsilon_{*,I}$ as high as the reddish bulge. The kinematics at larger radii would then be much better recovered by lower mass dark matter halo as found by \citet{SWELLSV}.

@@

@@

@@

\begin{table}
\centering
\begin{tabular}{cccccc}
\hline\hline
& & \multicolumn{2}{c}{Chabrier IMF} & \multicolumn{2}{c}{Salpeter IMF}\\
      &  $L$ [$10^{10}L_{\odot}$]                & $M_*$ [$10^{10}M_\odot$]               & $\Upsilon_*^\text{chab}$ & $M_*$ [$10^{10}M_\odot$] & $\Upsilon_*^\text{sal}$ \\\hline
bulge &   $3.10 \pm 0.15 $  & $7.8 \pm 1.8$ & $2.5 \pm 0.6$ & $14.5 \pm 3.7 $ & $4.7 \pm 1.2$ \\
disk  &   $2.35 \pm 0.11 $  & $2.9 \pm 0.7$ & $1.2 \pm 0.3$ & $5.2 \pm 1.1$ & $2.2 \pm 0.5$ \\
total &   $5.45 \pm 0.19$ & $10.6 \pm 1.9$& & $19.7 \pm 3.9$&\\\hline
\end{tabular}
\caption{Total I-band luminosity, stellar mass and mass-to-light ratio, calculated from the I-band AB magnitudes and stellar masses found for J133's bulge and disk by \citet{SWELLS} (their table 2) for comparison with this work. The transformation from AB magnitudes to the Johnson-Cousins I-Band used the relation $I[\text{mag}] = I[\text{ABmag}] - 0.309$ from \citet{FG1994} (their table 2). For the conversion from apparent magnitude to total luminosity the redshift $z=0.113$ \citet{SWELLSIII} was turned into a luminosity distance using the cosmology by \ref{WMAP5cosm}. }
\label{tab:previousresults}
\end{table}

>>>>

\subsection{Stellar mass-to-light ratios} \label{sec:MLdiscussion}

Further knowledge of the stellar populations within J1331 and their initial mass functions (IMF) would be useful. In particular a sophisticated guess for the \emph{stellar} mass-to-light ratio in J1331's bulge could be compared to our very reliable measurement of the \emph{total} mass-to-light ratio inside the Einstein radius (see Table \ref{tab:einsteinML} ) to support or contradict the presence of a lot of dark matter in the bulge. Traditional choices for the IMF are the bottom-heavy IMF by \citet{Salpeter1955},

@@

\\We compare our findings of the stellar mass-to-light ratios with the study by \cite{SWELLSV}: \cite{SWELLSV} found that the bulge of J1331 has an IMF more bottom-heavy than the Salpeter IMF from fitting a NFW halo to (1) the Einstein mass and (2) gas kinematics at larger radii $\gtrsim 8$ arcsec. We fitted the stellar kinematics within $\sim 3.5$ arcsec and the Einstein mass with a NFW halo and found $\Upsilon_\text{I,*}^\text{dyn} = 4.2 \pm 0.2$ (see Table \ref{tab:modelB4_bestfit}). \cite{SWELLSV}  found systematically lower NFW halos ($v_\text{circ}(5\text{arcsec}) \sim 120 \text{km s}^{-1}$ according to their Figure 2) than we did ($v_\text{circ}(5\text{arcsec}) \sim 200 \text{km s}^{-1}$, see Figure \ref{fig:kinematics}). [TO DO: Continue writing notes into text]

@@

\item We found an Einstein mass-to-light ratio of $\Upsilon_\text{I,tot}^\text{ein} = 5.6$ (see Table \ref{tab:einsteinML}). In case of a lower mass NFW halo, the largest contribution to this mass-to-light ratio would then be stellar mass. And the corresponding IMF would be more bottom heavy than Salpeter as well, like what \citet{SWELLSV} found.
\item The circular velocity curve $v_\text{circ}(R) = \left. sqrt{R \frac{\partial \Phi}{\partial R}} \right|_{z=0}$ \citet{SWELLSV} find for J1331 peaks at $v_\text{circ}(R\sim0.9 \text{arcsec}) \sim 350 \text{km s}^{-1}$ (from lensing constraints only) and at $v_\text{circ}(R\sim0.9 \text{arcsec}) \sim 370 \text{km s}^{-1}$ (when including a dark matter halo and gas kinematics in the fit). We find a very similar circular velocity curve (at least at small radii) for a mass-follows-light model with constant $\Upsilon_\text{I,tot}^\text{ein} = 5.6$, which peaks at $v_\text{circ}(R\sim0.9 \text{arcsec}) \sim 360 \text{km s}^{-1}$.

@@

\item The above discussion motivates the following speculation: In the absence of a strong dark matter contribution in the center, the overall stellar-mass-to-light ratio within the Einstein radius indicates a bottom-heavy Salpeter-like IMF, consistent with estimates from the velocity dispersion. At the same time the central dip can then be only explained, if there was an increase in stellar mass-to-light ratio with radius \textit{within} the bulge. \emph{If} there was a stellar population in the very center with an IMF close to the \citet{Chabrier2003} IMF, surrounded by a more bottom-heavy population, the central stellar kinematics would be well explained and be fully consistent with the lensing results, as found in Figure \ref{fig:modelG} and the surrounding text. It is not expected that the bluish disk of J1331 has a $\Upsilon_\text{I,*}$ as high as the reddish bulge. The kinematics at larger radii would then be much better recovered by lower mass dark matter halo as found by \citet{SWELLSV}.

@@

\item According to Figure 5c [TO DO: REF] the lensing result might not predict a mass to light ratio gradient inside Rein, but then again the mass slope $alpha=1$ was only weakly constrained.

@@

[TO DO: Einheitlich: $Upsilon_{I,*}^{comment}$, $\Upsilon_\text{I,tot}^\text{ein}$]

@@

\begin{table}
\centering
\begin{tabular}{cccccc}
\hline\hline
& & \multicolumn{2}{c}{Chabrier IMF} & \multicolumn{2}{c}{Salpeter IMF}\\
      &  $L$ [$10^{10}L_{\odot}$]                & $M_*$ [$10^{10}M_\odot$]               & $\Upsilon_\text{I,*}^\text{chab}$ & $M_*$ [$10^{10}M_\odot$] & $\Upsilon_\text{I,*}^\text{Salpeter}$ \\\hline
bulge &   $3.10 \pm 0.15 $  & $7.8 \pm 1.8$ & $2.5 \pm 0.6$ & $14.5 \pm 3.7 $ & $4.7 \pm 1.2$ \\
disk  &   $2.35 \pm 0.11 $  & $2.9 \pm 0.7$ & $1.2 \pm 0.3$ & $5.2 \pm 1.1$ & $2.2 \pm 0.5$ \\
total &   $5.45 \pm 0.19$ & $10.6 \pm 1.9$& & $19.7 \pm 3.9$&\\\hline
\end{tabular}
\caption{Total I-band luminosity, stellar mass and mass-to-light ratio, calculated from the I-band AB magnitudes and stellar masses found for J133's bulge and disk by \citet{SWELLS} (their table 2) for comparison with this work. The transformation from AB magnitudes to the Johnson-Cousins I-Band used the relation $I[\text{mag}] = I[\text{ABmag}] - 0.309$ from \citet{FG1994} (their table 2). For the conversion from apparent magnitude to total luminosity the redshift $z=0.113$ \citet{SWELLSIII} was turned into a luminosity distance using the cosmology by \ref{WMAP5cosm}. }
\label{tab:previousresults}
\end{table}

### 2015-09-04 docs/results_JAMwithSB_v1.01.tex
Finished section on mass / light comparison.

\paragraph{... with "Mass-follows-Light" and Velocity Anisotropy.} Our second JAM model is a mass-follows-light model, which are often used in dynamical JAM modelling (e.g. \citet{GlennEC,Cap06}), where the mass distribution ins generated by multiplying the light distribution in Table \ref{tab:MGEF814W} by a constant total mass-to-light ratio  $\Upsilon_{I,\text{tot}}^\text{dyn}$. This assumes that the dark matter is always a constant fraction of the total matter distribution everywhere. This simplified model sometimes gives good representations of the inner parts of galaxies, where the stellar component dominates. In addition to the free fit parameter  $\Upsilon_{I,\text{tot}}^\text{dyn}$, we also allow for a overall constant but non-zero velocity anisotropy $\beta_z$. The best fit is found by minimizing $\chi^2$ between the $v_\text{rms}$ data and model prediction and is demonstrated in Figure \ref{fig:JAM_modelA2}. For $\beta_z$ we impose the fitting limits $\beta_z \in [-0.5,+0.5]$. While the outer parts of galaxies often show radially biased velocity anisotropy up to $\sim 0.5$ (from dynamical modelling of observed elliptical galaxies (e.g. \citet{Kronawitter2000}) and cosmological simulations (e.g. \citet{2004MNRAS.352..535D,2001ApJ...557..533F}), the centers of galaxies are near-isotropic or have  negative velocity anisotropy \citep{2003ApJ...583...92G}. Only in extreme models (e.g. around in-spiralling supermassive black holes \citep{1997NewA....2..533Q}) velocity anisotropies as low as $\sim -1$ have been found. A lower limit of $\beta_z \geq -0.5$ is a realistic assumption for J1331, for which we do not expect extreme dynamical conditions. The best fit in Figure \ref{fig:JAM_modelA2} however strives to very negative velocity anisotropies to be able to get the deep central dip in the  $v_\text{rms}$ curve. But $\beta_z = -0.5$ is not even a remotely agreeable fit and lower anisotropies are not to be expected and realistic. We also tested radial profiles for $\beta_z(R)$ of the form proposed by \citet{BaesVanHese}, which was however equally unable to reproduce the data. We conclude, that this is due to the well-known degeneracy between anisotropy and mass profile [TO DO: REF] and the mass-follows-light model is \emph{not} a good representation of the mass distribution in J1331's inner regions.

>>>>

\paragraph{... with "Mass-follows-Light" and Velocity Anisotropy.} Our second JAM model is a mass-follows-light model, which are often used in dynamical JAM modelling (e.g. \citet{GlennEC,Cap06}), where the mass distribution ins generated by multiplying the light distribution in Table \ref{tab:MGEF814W} by a constant total mass-to-light ratio  $\Upsilon_{I,\text{tot}}^\text{dyn}$. This assumes that the dark matter is always a constant fraction of the total matter distribution everywhere. This simplified model sometimes gives good representations of the inner parts of galaxies, where the stellar component dominates. Testing this model for J1331 is also motivated by our findings from lensing in Section \ref{sec:results_lensing}, where around the Einstein radius total and luminous mass seem to have a similar distribution.
\\In addition to the free fit parameter  $\Upsilon_{I,\text{tot}}^\text{dyn}$, we also allow for a overall constant but non-zero velocity anisotropy $\beta_z$. The best fit is found by minimizing $\chi^2$ between the $v_\text{rms}$ data and model prediction and is demonstrated in Figure \ref{fig:JAM_modelA2}. For $\beta_z$ we impose the fitting limits $\beta_z \in [-0.5,+0.5]$. While the outer parts of galaxies often show radially biased velocity anisotropy up to $\sim 0.5$ (from dynamical modelling of observed elliptical galaxies (e.g. \citet{Kronawitter2000}) and cosmological simulations (e.g. \citet{2004MNRAS.352..535D,2001ApJ...557..533F}), the centers of galaxies are near-isotropic or have  negative velocity anisotropy \citep{2003ApJ...583...92G}. Only in extreme models (e.g. around in-spiralling supermassive black holes \citep{1997NewA....2..533Q}) velocity anisotropies as low as $\sim -1$ have been found. A lower limit of $\beta_z \geq -0.5$ is a realistic assumption for J1331, for which we do not expect extreme dynamical conditions. The best fit in Figure \ref{fig:JAM_modelA2} however strives to very negative velocity anisotropies to be able to get the deep central dip in the  $v_\text{rms}$ curve. But $\beta_z = -0.5$ is not even a remotely agreeable fit and lower anisotropies are not to be expected and realistic. We also tested radial profiles for $\beta_z(R)$ of the form proposed by \citet{BaesVanHese}, which was however equally unable to reproduce the data. We conclude, that this is due to the well-known degeneracy between anisotropy and mass profile [TO DO: REF] and the mass-follows-light model is \emph{not} a good representation of the mass distribution in J1331's inner regions.
